Comment: Reasoning about Interference Between Units

### 2015-12-21 paper/introduction.tex
Working on finishing the simplified discussion part

Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions. In the case where we know
the fixed adjacency matrix of the network, $\bS$, and where we imagine that
network attributes (like degree) of a node play a role in the mechanism by
which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

@@

To provide a little intuition about the SSR test statistic as compared to the
KS test statistic,

we created a small simulation study of 256 units with half
assigned to a binary treatment completely at random. We compared two outcomes
--- a Normal outcome and a skewed outcome, both with mean of 10 and standard
deviation of 1. And we compared two models of effects --- a constant additive
effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
the truth to be the sharp null of no effects $\tau=0$ for the additive model
and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e. the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

>>>>

Fisher/Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions: For example, the KS test
  uses the maximum difference in the empirical cumulative distributions of
  each treatment group calculated without regard for the relationship between
the treated and control distributions, thereby ignoring information that could
increase the precision of the test. The simplest version of the SSR test
statistic merely sums the difference between the mean of the outcome implied
by the hypothesis and individual outcomes:

\begin{equation}
 {\T(\yu,\bz)}_{\text{SSR}} \equiv \sum_{i} {( \widetilde y_{i,Z_i=0} - \bar{\widetilde{y}}_{i,Z_i=0} )}^2 \label{eq:ssr1}
\end{equation}

To provide a little intuition about the SSR test statistic as compared to the
KS test statistic, we created a small simulation study of 256 units with half
assigned to a binary treatment completely at random. We compared the
performance of the two test statistics on two outcomes with no interference
--- a Normal outcome and a skewed outcome, both with mean of 10 and standard
deviation of 1 as shown in Figure~\ref{fig:simpleoutcomes}.

\begin{figure}[H]\centering
  \includegraphics[width=.5\textwidth]{figures/ksvsssroutcomes.pdf}
  \caption{}\label{fig:simpleoutcomes}
\end{figure}

And we compared two models of effects --- a constant additive
effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
the truth to be the sharp null of no effects $\tau=0$ for the additive model
and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e. the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

In the case where we know
the fixed adjacency matrix of the network, $\bS$, and where we imagine that
network attributes (like degree) of a node play a role in the mechanism by
which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

@@
